Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-07-23 manuscript/response_to_reviews.Rmd

**As stated above, we have removed the $Rt$ criterion from the drought identification process, and now present this information for the first time in this section, adding the following statement: "Across the entire study period (1950-2009), the focal drought years were the three years with the largest fraction of trees exhibiting $Rt \le 0.7$."**

- Line 258: so this means that they are no sensitive to the drought? Which trees, any species in particular? Is this a consequence of stand dynamics shadowing the impact of drought events?

**It's not unprecedented to see some trees exhibit increased growth during drought, and not surprising. Stand dynamics and other factors result in a lot of scatter in the relationship between tree growth and climate, so while the distribution of $Rt$ in Fig. 1b was shifted left (normal year should be normally distributed around 1), there still are some trees with increased growth. Our analysis helps to work out which trees these were (often shorter trees).**

**We address this question in the discussion, where we have added a statement about how this is not unprecedented: "Across all three droughts, the majority of trees experienced reduced growth, but a substantial portion had increased growth (Fig. 1b), consistent with prior observations that smaller trees can exhibit increased growth rates during drought (Bennett et al. 2015). It is likely because of the moderate impact of these droughts, along with other factors influencing tree growth (e.g., stand dynamics), that our best models characterize only a modest amount of variation"**

@@

**Unless the reviewers and editor strongly disagree, we would request to keep this figure to clarify the main manuscript. We have revised the manuscript such that this figure is better integrated with the hypothesis testing (Table 1) and interpretation of results. We agree that similar patterns are common to any forest, and are not trying to claim this as any kind of novel result, but we're not aware of studies showing this depiction of height profiles for a temperate forest. (Such patterns are not shown in some relevant textbooks: Gordan Bonan's 2016 *Ecological Climatology* and Norman & Campbell's *Environmental Biophysics*. If we're missing studies that show such vertical profiles, we'd appreciate the references.)**

**We did create two new figures showing the results of the tree-ring analysis (and moved two tables to the SI).**

- Line 330-331: any evidence to support this? It looks rather speculative. As suggested above, does any species particularly show positive Rt or positive Rt are randomly distributed in your trees?

*Current text, with phrase of concern in bold: "Across all three droughts, the majority of trees experienced reduced growth, but a substantial portion had increased growth (Fig. 1b), consistent with prior observations that smaller trees can exhibit increased growth rates during drought [@bennett_larger_2015]." *

### 2020-07-23 manuscript/climate_sensitivity_draft.Rmd
removed claim lacking ref

The droughts considered here were of a magnitude that has occurred with an average frequency of approximately once every 10-15 years [Fig. **1a**, @helcoski_growing_2019] and had substantial but not devastating impacts on tree growth (Figs. **1b**). These droughts were classified as severe ($PDSI$ < -3.0; 1977) or extreme ($PDSI$ < -4.0; 1966, 1999) at our site and have been linked to tree mortality in the eastern United States [@druckenbrod_redefining_2019]. However, extreme, multiannual droughts such as the so-called “megadroughts” of this type that have triggered massive tree die-off in other regions [e.g., @allen_global_2010; @stovall_tree_2019] have not occurred in the Eastern United States within the past several decades [@clark_impacts_2016]. Of the droughts considered here, the 1966 drought, which was preceded by two years of dry conditions (Fig. **S1**), severely stressed a larger portion of trees (Fig. **1b**). The tendency for large trees to have lowest resistance was most pronounced in this drought, consistent with other findings that this physiological response increases with drought severity [@bennett_larger_2015; @stovall_tree_2019]. Across all three droughts, the majority of trees experienced reduced growth, but a substantial portion had increased growth (Fig. **1b**), potentially due to decreased leaf area of competitors during the drought (**REF--if we can find one**), and consistent with prior observations that smaller trees can exhibit increased growth rates during drought [@bennett_larger_2015]. It is likely because of the moderate impact of these droughts, along with other factors influencing tree growth (e.g., stand dynamics), that our best models characterize only a modest amount of variation in $Rt$: 11-12% for all droughts combined, and 18-25% for each individual drought (Fig. **S6**; Table S6).

>>>>

The droughts considered here were of a magnitude that has occurred with an average frequency of approximately once every 10-15 years [Fig. **1a**, @helcoski_growing_2019] and had substantial but not devastating impacts on tree growth (Figs. **1b**). These droughts were classified as severe ($PDSI$ < -3.0; 1977) or extreme ($PDSI$ < -4.0; 1966, 1999) at our site and have been linked to tree mortality in the eastern United States [@druckenbrod_redefining_2019]. However, extreme, multiannual droughts such as the so-called “megadroughts” of this type that have triggered massive tree die-off in other regions [e.g., @allen_global_2010; @stovall_tree_2019] have not occurred in the Eastern United States within the past several decades [@clark_impacts_2016]. Of the droughts considered here, the 1966 drought, which was preceded by two years of dry conditions (Fig. **S1**), severely stressed a larger portion of trees (Fig. **1b**). The tendency for large trees to have lowest resistance was most pronounced in this drought, consistent with other findings that this physiological response increases with drought severity [@bennett_larger_2015; @stovall_tree_2019]. Across all three droughts, the majority of trees experienced reduced growth, but a substantial portion had increased growth (Fig. **1b**), consistent with prior observations that smaller trees can exhibit increased growth rates during drought [@bennett_larger_2015]. It is likely because of the moderate impact of these droughts, along with other factors influencing tree growth (e.g., stand dynamics), that our best models characterize only a modest amount of variation in $Rt$: 11-12% for all droughts combined, and 18-25% for each individual drought (Fig. **S6**; Table S6).
